---
title: Analyse overlap signatures
output:
  html_notebook: default
  html_document: default
  pdf_document: default
---
```{r, include = FALSE}
library(GenomicRanges)
library(stringr)
library(dplyr)
library(tidyr)
library(tidyverse)
"%ni%" <- Negate("%in%")
```


```{r}
# setwd("~/Github/Ting_signature_analyses/")

annotation_methylation_CpGs <- read.csv("BPmetCan.hg19.csv", stringsAsFactors = F)
annotation_methylation_CpGs$seqnames <- substr(annotation_methylation_CpGs$seqnames, 4, 6)

load("pchic.RData")
pchic <- data.frame(pchic[rowSums(pchic[,11:20] >= 0) >= 1, 1:10]) %>% na.omit(.)
List_Promoter <- unique(paste(pchic$baitChr, pchic$baitStart, sep = "_"))
colnames(pchic)[c(1:5, 6:10)] <- rep(c("chr", "start", "end", "ID", "Name"), 2)
PCHiC_bed <- unique(rbind(pchic[, c(1:3, 5)], pchic[, c(6:8, 10)]))
PCHiC_GRange <- GRanges(
  seqnames = PCHiC_bed$chr,
  IRanges(start = PCHiC_bed$start, end = PCHiC_bed$end),
  Gene_Pchic = PCHiC_bed$Name,
  start_fragment = PCHiC_bed$start,
  end_fragment = PCHiC_bed$end
)
PCHiC_GRange$ID <- paste(PCHiC_bed$chr, PCHiC_bed$start, sep = "_")
colnames(pchic) <- c("chr_bait", "start_bait", "end_bait", "ID_bait", "Name_bait", "chr_oe", "start_oe", "end_oe", "ID_oe", "Name_oe")
pchic$IDbait <- paste(pchic$chr_bait, pchic$start_bait, sep = "_")
pchic$IDoe <- paste(pchic$chr_oe, pchic$start_oe, sep = "_")

Blueprint_network <- read.csv("BLUEPRINT_fragments_good.tsv", sep = "\t", stringsAsFactors = F) %>%
  dplyr::select(., "chr", "start", "end", "type", "ensembl", "gene_names", "intronic_regions", "type") %>% 
  separate_rows(., gene_names, sep = " ")

Final.Tumor.RNAseq.sig <- read.delim("BPRNACan_150.txt", stringsAsFactors = F) %>% 
  .[, "Gene"]

Blueprint_Granges <- GRanges(
  seqnames = Blueprint_network$chr,
  ranges = IRanges(start = Blueprint_network$start, end = Blueprint_network$end),
  Gene_name = Blueprint_network$gene_names,
  Promoter = Blueprint_network$type
)
```


Overlapping between Chromatin fragment from PCHiC data and Blueprint data
```{r}
overlaps_Blueprint_pchic <- findOverlaps(PCHiC_GRange, Blueprint_Granges)
# partie metadata du granges en data frame, puis merge des deux metadata de phic_grange et BP_grange
match_hit_BP_Pchic <- data.frame(mcols(PCHiC_GRange[queryHits(overlaps_Blueprint_pchic), ]),
                                 data.frame(mcols(Blueprint_Granges[subjectHits(overlaps_Blueprint_pchic), ])),
                                 stringsAsFactors = T)

message("=== Number of nodes in BP network ===")
Nodes_in_BP_network <- unique(match_hit_BP_Pchic$ID)
length(Nodes_in_BP_network)

message("== Number of promoters nodes in BP network ===")
match_hit_BP_promoter_pchic <- match_hit_BP_Pchic[match_hit_BP_Pchic$Promoter == "P",]
Nodes_in_BP_Promoter <- unique(match_hit_BP_promoter_pchic$ID)
length(Nodes_in_BP_Promoter)

Genes_BP <- unique(match_hit_BP_Pchic$Gene_name)
Genes_BP_Promoter_signature <- Genes_BP[Genes_BP %in% Final.Tumor.RNAseq.sig]

message("== Number of Promoter in BPRNACan signature ==")
length(Genes_BP_Promoter_signature)
```

Overlapping between CpGs from BPmetCan signature and PCHiC data
```{r}
CpGs_GRanges <- GRanges(
  seqnames = annotation_methylation_CpGs$seqnames,
  ranges = IRanges(start = annotation_methylation_CpGs$start, end = annotation_methylation_CpGs$end),
  chr_cpg = annotation_methylation_CpGs$seqnames
)

overlaps_CpGs_pchic <- findOverlaps(PCHiC_GRange, CpGs_GRanges)
match_hit_CpGs_Pchic <- data.frame(mcols(PCHiC_GRange[queryHits(overlaps_CpGs_pchic),]), data.frame(mcols(CpGs_GRanges[subjectHits(overlaps_CpGs_pchic),])))

message("=== Number of CpGs from BPmetCan matching a node in the network ===")
Nodes_in_CpG <- unique(match_hit_CpGs_Pchic$ID)
length(Nodes_in_CpG)
```

Overlapping between the overlapped CpGs in PCHiC data and Blueprint data
```{r}
CpGs_pchic_GRanges <- GRanges(
  seqnames = match_hit_CpGs_Pchic$chr_cpg,
  ranges = IRanges(start = match_hit_CpGs_Pchic$start_fragment, end = match_hit_CpGs_Pchic$end_fragment),
  chr_cpg = match_hit_CpGs_Pchic$chr_cpg,
  ID = match_hit_CpGs_Pchic$ID
)

overlaps_CpGs_Blueprint <- findOverlaps(Blueprint_Granges, CpGs_pchic_GRanges)
match_hit_CpGs_Blueprint <- data.frame(mcols(Blueprint_Granges[queryHits(overlaps_CpGs_Blueprint),]),
                                       data.frame(mcols(CpGs_pchic_GRanges[subjectHits(overlaps_CpGs_Blueprint),])))

message("=== CpGs nodes associated to a promoter in blueprint data ===")
CpGs_nodes_promoter <- unique(match_hit_CpGs_Blueprint[match_hit_CpGs_Blueprint$Promoter == "P", "ID"])
length(CpGs_nodes_promoter)
message("=== Genes from promoter blueprint network associated to CpGsnodes from sig ===")
CpGs_promoter <- unique(match_hit_CpGs_Blueprint[match_hit_CpGs_Blueprint$Promoter == "P", "Gene_name"])
length(CpGs_promoter)
message("=== CPGsnodes that are NOT associated with a promoter from blueprint network ===")
CpGs_not_promoter <- unique(match_hit_CpGs_Blueprint[match_hit_CpGs_Blueprint$Promoter == "O", "ID"])
length(CpGs_not_promoter)


```

Promoters from the overlaps between PCHiC blueprint and CpGs
```{r}
Nodes_with_CpGs_sig_and_gene_sig <- match_hit_CpGs_Blueprint[match_hit_CpGs_Blueprint$Gene_name %in% Final.Tumor.RNAseq.sig & match_hit_CpGs_Blueprint$Promoter == "P", "Gene_name"]
message("=== Number of BPmetCan promoter-associated CpGs nodes that are in gene list from BPRNA signature ===")
length(Nodes_with_CpGs_sig_and_gene_sig)

message("=== Number of genes from CpGs sig that are in gene list from BPRNA signature ===")
Genes_sig_in_CpG_sig <- unique(Nodes_with_CpGs_sig_and_gene_sig)
length(Genes_sig_in_CpG_sig)
```

Neighborhood of the CpGs signatures
```{r}
Neighbor_network <- unique(rbind(pchic[pchic$IDbait %in% match_hit_CpGs_Pchic$ID,],pchic[pchic$IDoe %in% match_hit_CpGs_Pchic$ID,]))[,c(1:3,11, 5:8,12, 10)]

colnames(Neighbor_network) <- rep(c("chr", "start", "end", "ID", "Gene_name"), 2)
Neighbor_nodes <- unique(rbind(Neighbor_network[,c(1:5)], Neighbor_network[,c(6:10)]))

Promoter_neighbors <- match_hit_BP_promoter_pchic[match_hit_BP_promoter_pchic$ID %in% Neighbor_nodes$ID, ]
message("=== Number of neighbor nodes of CpGs sig in promoter ===")
Nodes_promoter_neighbor_CpG_sig <- unique(Promoter_neighbors$ID)
length(Nodes_promoter_neighbor_CpG_sig)

message("=== Number of promoter of genes in neighbor CpGs sig ===")
Promoter_neighbors_genes <- unique(Promoter_neighbors$Gene_name)
length(Promoter_neighbors_genes)
```
Neighborhood of the CpGs signatures not in promoter

```{r}
Neighbor_not_prom_network <- unique(rbind(pchic[pchic$IDbait %in% CpGs_not_promoter,],pchic[pchic$IDoe %in% CpGs_not_promoter,]))[,c(1:3,11, 5:8,12, 10)]

colnames(Neighbor_not_prom_network) <- rep(c("chr", "start", "end", "ID", "Gene_name"), 2)
Neighbor_not_prom_nodes <- unique(rbind(Neighbor_not_prom_network[,c(1:5)], Neighbor_not_prom_network[,c(6:10)]))

Promoter_neighbors_of_not_prom <- match_hit_BP_promoter_pchic[match_hit_BP_promoter_pchic$ID %in% Neighbor_not_prom_nodes$ID, ]
message("=== Number of neighbor nodes promoter from non promoter Cpg ===")
Nodes_neighbor_non_prom_CpGs_sig <- unique(Promoter_neighbors_of_not_prom$ID)
length(Nodes_neighbor_non_prom_CpGs_sig)

message("=== Number of genes in neighbor CpGs sig non promoter ===")
Promoter_neighbors_of_not_prom_genes <- unique(Promoter_neighbors_of_not_prom$Gene_name)
length(Promoter_neighbors_of_not_prom_genes)
message("=== Number of non sig genes in the neighborhood of CpGs sig non promoter ===")
Non_sig_genes_CpGs_neighbor <- Promoter_neighbors_of_not_prom_genes[Promoter_neighbors_of_not_prom_genes %ni% Final.Tumor.RNAseq.sig]
length(Non_sig_genes_CpGs_neighbor)

```

Neighborhood of the CpGs sig in promoter of Genes from the signature

```{r}
Nodes_intersect <- match_hit_CpGs_Blueprint[match_hit_CpGs_Blueprint$Promoter == "P" & match_hit_CpGs_Blueprint$Gene_name %in% Final.Tumor.RNAseq.sig, "ID"]
Neighbor_intersect_signature <- unique(rbind(pchic[pchic$IDbait %in% Nodes_intersect,],pchic[pchic$IDoe %in% Nodes_intersect,]))[,c(1:3,11, 5:8,12, 10)]

colnames(Neighbor_intersect_signature) <- rep(c("chr", "start", "end", "ID", "Gene_name"), 2)
Neighbor_intersect_signature_nodes <- unique(rbind(Neighbor_intersect_signature[,c(1:5)], Neighbor_intersect_signature[,c(6:10)]))

Promoter_neighbor_intersect_signature <- match_hit_BP_promoter_pchic[match_hit_BP_promoter_pchic$ID %in% Neighbor_intersect_signature_nodes$ID, ]
message("=== Number of neighbor nodes of the CpGs sig in promoter of Genes sig ===")
Nodes_neigbhor_intersection_CpGs_genes_sig <- unique(Promoter_neighbor_intersect_signature$ID)
length(Nodes_neigbhor_intersection_CpGs_genes_sig)

message("=== Number of promoter genes in neighbor CpGs sig of Genes sig ===")
Promoter_neighbor_intersect_signature <- unique(Promoter_neighbor_intersect_signature$Gene_name)
length(Promoter_neighbor_intersect_signature)

message("=== Number of non sig genes in the neighborhood of CpGs sig and Genes sig ===")
Non_sig_genes_CpGs_Genes_neighbor <- Promoter_neighbor_intersect_signature[Promoter_neighbor_intersect_signature %ni% Final.Tumor.RNAseq.sig]
length(Non_sig_genes_CpGs_Genes_neighbor)
```


Promoter that are both in Gene sig and neighbor of CpGs sig
```{r}
Promoter_neighbors_genes_sig <- Promoter_neighbors_genes[Promoter_neighbors_genes %in% Final.Tumor.RNAseq.sig]
message("=== Number of promoters from BPRNA signature genes in neighborhood of CpGs sig ===")
length(unique(Promoter_neighbors_genes_sig))
```

Promoter
```{r}
unik_CpGs_promoter <- unique(CpGs_promoter)
solo_genes <- unik_CpGs_promoter[unik_CpGs_promoter %ni% Final.Tumor.RNAseq.sig]
message("=== Genes associated with a promoter in BPmetCan, that are not present in BPRNACan ===")
length(solo_genes)
```

## GO on CPGs related genes
```{r}
genes_for_GO <- unique( unlist( str_split(Promoter_neighbors_genes, " ") ) )
write.table(genes_for_GO, "cpgs_genes.txt", row.names = F, col.names = F, quote = F)
# launch on pantherDB http://geneontology.org
GO <-read.table("analysis.txt", sep="\t", header=T, skip = 11)[,c(1,7)]
GO$GO.biological.process.complete <- str_extract(GO$GO.biological.process.complete, "GO:\\d+")
write.table(GO, "go_promoter_genes.txt", row.names = F, col.names = F, quote = F)
# Launch on REVIGO http://revigo.irb.hr
```

```{r}
sessionInfo()
```